The Talking Heads experiment

### 2015-04-29 chap8/chap8.tex
part 2 of martin haspelmath's comments

\section{The Laboratorum exhibition}

The Laboratorium\is{Laboratorium exhibition}
exhibition was a major artistic event in Europe during the summer of 1999 organised by 
Bruno Verbergt of an organisation called \smplenquote{Antwerpen Open}.
It was one of the first exhibitions that put so much emphasis on art as a research 
activity and on profound interactions between art and science.
Participants included both scientists and science historians such as Peter Galison, Bruno Latour, Israel Rosenfield
and Isabelle Stengers, as well as architects such as Rem Koolhaas and artists with an affinity for research 
such as Peter Fischli and David Weiss, Gabriel Orozco, Carsten H\"oller, Panamarenko, Lawrence Weiner and others. 
The exhibition was visited by 8000 people.

@@

The installation in Antwerp was announced as a ``Laboratory for Cognitive Robotics and Teleportation" \citep{Steels:99b}.
It featured various rooms as shown in the layout in \figref{fig:layout}. The rooms had different functions: 
\begin{enumerate} 
\item The {\itshape central room} which was visible on entering the space contained the experiment itself, i.e. the 
two cameras mounted on tripods, the computers driving them, the screen with geometric figures, and a projection of 
what went on. The activities of the agents were audible through a narration. 
\clearpage
\item A {\itshape reading room} contained background philosophical and scientific 
papers and excerpts from books about language, meaning and their origins. 
\item A {\itshape user interface room} contained a workstation where visitors to the exhibition could create their own 
agents, direct them to play games on certain sites, and teach new words to their agents. 
\end{enumerate}

@@

A typical example of the kind of images that the cameras picked up are shown in \figref{fig:images}. 
To play a game, the agent had to find a sufficiently delineated group of objects and each object had 
to exceed a minimal size. Initially quite complex configurations were put on the white board but this made it very difficult for agents to find a coherent group and to develop good concepts for referring to the topic. The right 
image in \figref{fig:images} provides clear examples of up, down, left and right, different colors, and 
also opportunities to use multiple words (such as \smplenquote{red bottom}).

@@

A typical example of an interaction is shown in \figref{fig:interaction}. 
The speaker (called ``rubber", an agent created and named by a user) has selected the blue object at the bottom left 
of the screen. We see the discrimination trees of the agent at that point and the data for each of the objects 
recognised. The coordinates of the respective objects (scaled within the coordinate system of 
the captured image) are (0.0,1.0) for the blue object, (1.0, 0.96) for the red one, and (0.42, 0.0) for the green one. 
The values are displayed as \textsc{hp} (horizontal position), \textsc{vp} (vertical position), \textsc{h} (height), \textsc{w} (width), \textsc{a} (area), 
\textsc{r} (red), \textsc{g} (green), \textsc{y} (yellow), \textsc{b} (blue), \textsc{l} (lightness). 
The blue object is chosen as the topic and the value on the blue channel has the most discriminative power, although 
there are other possibilities. There are three competing words for naming blue: \smplenquote{Xagadude}, \smplenquote{Nibidesu}, and \smplenquote{Tetipi}. 
None of them have a score higher than 0 and so a random choice of \smplenquote{Xagadude} is made.

@@

An example of how the web interface displayed a single game is shown in \figref{fig:thwebsite-agent}. 
There are two agents: Antonusius is the speaker and Zelebot is the hearer. 
The green object has been chosen as topic and correctly 
recognised by the hearer using the word \smplenquote{kazozo}. The meaning of the word was not visible through the interface because we 
wanted users to learn themselves the language of the agents.

\begin{figure}[htb]
  \centerline{\includegraphics[width=.85\textwidth]{chap8/figures/zelebot.pdf}}
\caption{\label{fig:thwebsite-agent}Example of an interaction that took part in Antwerp. This game was a success. The agents recognised the same
set of objects and the meaning of \smplenquote{kazozo} was effective in finding the same topic. 
}
\end{figure}

@@

The first \dblenquote{Talking Heads} experiment ran for 4 months during the summer of 1999 and showed the validity of the mechanisms that were used for the agent architecture and of the interaction patterns and group dynamics of the agents. A shared lexicon and an underlying conceptual repertoire emerged, enabling successful communication by the agents about the scenes before them. In total, 400,000 grounded games were played. The population of agents rose to just under 2000, increasing steadily over the period of the experiment. Despite the many perturbations due to grounding, intermittent technical failures, a continuous influx of new agents entering the population, and unpredictable human interaction, the lexicon was maintained throughout this period.

@@

\figref{fig:droite} shows another example of the semiotic dynamics in the experiment, this time looking at all the 
meanings for a particular word, namely \smplenquote{droite} (meaning \smplenquote{to the right} in French). This word clearly has been introduced
by a human user and the dominating meaning progressively becomes a region (between 0.75 and 1.0) on the horizontal position, 
as could be expected. On the way we see some confusion, Specifically there must be objects that appear both on the right 
and up (vertical position between 0.5 and 1.0).

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap8/figures/droite.pdf}}
\caption{\label{fig:droite}Form-Meaning diagram: Different meanings associated with the word \smplenquote{droite}. The dominant meaning is 
in line with the human use of the word, namely horizontal position (\textsc{hpos}) to the right of the image.}
\end{figure}

\figref{fig:bozopite} shows different meanings competing for the same word "bozopite". Dominant ones are 
a large area (\textsc{area} scaled between 0.5 and 1.0) or extended width (\textsc{width} scaled between 0.5 and 1.0). These two concepts
apparently could both be applied in many situations, but at some point new situations appeared that 
caused a symmetry breaking and \textsc{area} became dominant.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap8/figures/bozopite.pdf}}
\caption{\label{fig:bozopite}Form-Meaning diagram: Different meanings associated with the word \smplenquote{bozopite}. 
}
\end{figure}

It was also possible that different meanings which were highly compatible were maintained in the population. 
An example is shown in the form-meaning diagram in \figref{fig:down}, which plots the average 
frequency of meanings for the word \smplenquote{down}. The meanings are all concepts on the vertical position channel 
\textsc{vpos}, but they carve out smaller and smaller regions.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap8/figures/down.pdf}}
\caption{\label{fig:down}Form-Meaning diagram: Different meanings associated with the word ``down". There are in particular two meanings. 
one is a region on the red-channel and another one a region on the vertical position channel. 
}
\end{figure}

@@

Numerous talks were given and various papers published in scientific fora. There were also various talks within 
the art context.\footnote{Examples are a \smplenquote{gallery talk} as Salon 3 at the Elephant and Castle Centre
in London on 2 December 1998, organised by Hans Ulrich Obrist and Molly Nesbit, and a presentation together with 
Hans-Ulrich Obrist and Rem Koolhaas in Antwerp to launch the Laboratorium book on 3 October 2001.}
The experiment received wide coverage in 
the media thanks to its public exposure as part of a major exhibition. The Süddeutsche Zeitung called it 
\dblenquote{Angels with Internet wings}. All this led to invitations to show the experiment
also in other venues, as discussed in the next chapter. 
Regrettably, the challenge just to keep the experiment in the air with the available human resources
and the pressure for going on with new experiments prevented us from doing more adequate data gathering 
and analysis. Nevertheless some analyses appeared, particularly as carried out by Frederic Kaplan (\citealt{Kaplan:2001}).

>>>>

\section{The Laboratorium exhibition}

The Laboratorium\is{Laboratorium exhibition}
exhibition was a major artistic event in Europe during the summer of 1999 organised by 
Bruno Verbergt of an organisation called \textit{Antwerpen Open}.
It was one of the first exhibitions that put so much emphasis on art as a research 
activity and on profound interactions between art and science.
Participants included both scientists and science historians such as Peter Galison, Bruno Latour, Israel Rosenfield
and Isabelle Stengers, as well as architects such as Rem Koolhaas and artists with an affinity for research 
such as Peter Fischli and David Weiss, Gabriel Orozco, Carsten H\"oller, Panamarenko, Lawrence Weiner and others. 
The exhibition was visited by 8000 people.

@@

The installation in Antwerp was announced as a ``Laboratory for Cognitive Robotics and Teleportation" \citep{Steels:99b}.
It featured various rooms as shown in the layout in \figref{fig:layout}. The rooms had different functions: 
\begin{enumerate} 
\item The {\scshape central room} which was visible on entering the space contained the experiment itself, i.e. the 
two cameras mounted on tripods, the computers driving them, the screen with geometric figures, and a projection of 
what went on. The activities of the agents were audible through a narration. 
\clearpage
\item A {\scshape reading room} contained background philosophical and scientific 
papers and excerpts from books about language, meaning and their origins. 
\item A {\scshape user interface room} contained a workstation where visitors to the exhibition could create their own 
agents, direct them to play games on certain sites, and teach new words to their agents. 
\end{enumerate}

@@

A typical example of the kind of images that the cameras picked up are shown in \figref{fig:images}. 
To play a game, the agent had to find a sufficiently delineated group of objects and each object had 
to exceed a minimal size. Initially quite complex configurations were put on the white board but this made it very difficult for agents to find a coherent group and to develop good concepts for referring to the topic. The right 
image in \figref{fig:images} provides clear examples of up, down, left and right, different colors, and 
also opportunities to use multiple words (such as \textit{red bottom}).

@@

A typical example of an interaction is shown in \figref{fig:interaction}. 
The speaker (called ``rubber", an agent created and named by a user) has selected the blue object at the bottom left 
of the screen. We see the discrimination trees of the agent at that point and the data for each of the objects 
recognised. The coordinates of the respective objects (scaled within the coordinate system of 
the captured image) are (0.0,1.0) for the blue object, (1.0, 0.96) for the red one, and (0.42, 0.0) for the green one. 
The values are displayed as \textsc{hp} (horizontal position), \textsc{vp} (vertical position), \textsc{h} (height), \textsc{w} (width), \textsc{a} (area), 
\textsc{r} (red), \textsc{g} (green), \textsc{y} (yellow), \textsc{b} (blue), \textsc{l} (lightness). 
The blue object is chosen as the topic and the value on the blue channel has the most discriminative power, although 
there are other possibilities. There are three competing words for naming blue: \textit{Xagadude}, \textit{Nibidesu}, and \textit{Tetipi}. 
None of them have a score higher than 0 and so a random choice of \textit{Xagadude} is made.

@@

An example of how the web interface displayed a single game is shown in \figref{fig:thwebsite-agent}. 
There are two agents: Antonusius is the speaker and Zelebot is the hearer. 
The green object has been chosen as topic and correctly 
recognised by the hearer using the word \textit{kazozo}. The meaning of the word was not visible through the interface because we 
wanted users to learn themselves the language of the agents.

\begin{figure}[htb]
  \centerline{\includegraphics[width=.85\textwidth]{chap8/figures/zelebot.pdf}}
\caption{\label{fig:thwebsite-agent}Example of an interaction that took part in Antwerp. This game was a success. The agents recognised the same
set of objects and the meaning of \textit{kazozo} was effective in finding the same topic. 
}
\end{figure}

@@

The first Talking Heads experiment ran for 4 months during the summer of 1999 and showed the validity of the mechanisms that were used for the agent architecture and of the interaction patterns and group dynamics of the agents. A shared lexicon and an underlying conceptual repertoire emerged, enabling successful communication by the agents about the scenes before them. In total, 400,000 grounded games were played. The population of agents rose to just under 2000, increasing steadily over the period of the experiment. Despite the many perturbations due to grounding, intermittent technical failures, a continuous influx of new agents entering the population, and unpredictable human interaction, the lexicon was maintained throughout this period.

@@

\figref{fig:droite} shows another example of the semiotic dynamics in the experiment, this time looking at all the 
meanings for a particular word, namely \textit{droite} (meaning \smplenquote{to the right} in French). This word clearly has been introduced
by a human user and the dominating meaning progressively becomes a region (between 0.75 and 1.0) on the horizontal position, 
as could be expected. On the way we see some confusion, Specifically there must be objects that appear both on the right 
and up (vertical position between 0.5 and 1.0).

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap8/figures/droite.pdf}}
\caption{\label{fig:droite}Form-Meaning diagram: Different meanings associated with the word \textit{droite}. The dominant meaning is 
in line with the human use of the word, namely horizontal position (\textsc{hpos}) to the right of the image.}
\end{figure}

\figref{fig:bozopite} shows different meanings competing for the same word \textit{bozopite}. Dominant ones are 
a large area (\textsc{area} scaled between 0.5 and 1.0) or extended width (\textsc{width} scaled between 0.5 and 1.0). These two concepts
apparently could both be applied in many situations, but at some point new situations appeared that 
caused a symmetry breaking and \textsc{area} became dominant.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap8/figures/bozopite.pdf}}
\caption{\label{fig:bozopite}Form-Meaning diagram: Different meanings associated with the word \textit{bozopite}. 
}
\end{figure}

It was also possible that different meanings which were highly compatible were maintained in the population. 
An example is shown in the form-meaning diagram in \figref{fig:down}, which plots the average 
frequency of meanings for the word \textit{down}. The meanings are all concepts on the vertical position channel 
\textsc{vpos}, but they carve out smaller and smaller regions.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap8/figures/down.pdf}}
\caption{\label{fig:down}Form-Meaning diagram: Different meanings associated with the word \textit{down}. There are in particular two meanings. 
one is a region on the red-channel and another one a region on the vertical position channel. 
}
\end{figure}

@@

Numerous talks were given and various papers published in scientific fora. There were also various talks within 
the art context.\footnote{Examples are a \dblenquote{gallery talk} as Salon 3 at the Elephant and Castle Centre
in London on 2 December 1998, organised by Hans Ulrich Obrist and Molly Nesbit, and a presentation together with 
Hans-Ulrich Obrist and Rem Koolhaas in Antwerp to launch the Laboratorium book on 3 October 2001.}
The experiment received wide coverage in 
the media thanks to its public exposure as part of a major exhibition. The \textit{Süddeutsche Zeitung} called it 
\dblenquote{Angels with Internet wings}. All this led to invitations to show the experiment
also in other venues, as discussed in the next chapter. 
Regrettably, the challenge just to keep the experiment in the air with the available human resources
and the pressure for going on with new experiments prevented us from doing more adequate data gathering 
and analysis. Nevertheless some analyses appeared, particularly as carried out by Frederic Kaplan (\citealt{Kaplan:2001}).

### 2015-04-29 chap9/chap9.tex
part 2 of martin haspelmath's comments

\noindent
{\bfshape Situated robots and teleports}

@@

\noindent
{\bfshape Constructing perceptually grounded concepts}

@@

\noindent
{\bfshape Sources of noise}

@@

\noindent
{\bfshape Cultural evolution}

@@

We pick up the dialog on the 25 February 2000. Until that date the forum was very active both 
with general discussions about the future of intelligence or the origins of language, and with specific questions, mostly 
why agents were so slow in playing games. Many users were too impatient and did not seem to realise that playing 
a language game could easily take a minute or two. However the major problem was the deviating alignment of the cameras, 
which had to guarantee that there was a common frame of reference between the agents, and hence the 
possibility of sharing attention. The cameras were mounted on tripods. When somebody accidentally 
moved the tripod, the frame of reference was no longer exact. As mentioned earlier, traffic (particularly subway 
traffic in London) caused vibrations of the cameras so that they kept shifting and getting out of balance.

@@

{\bfshape 2000-03-03 20:08:36 schedski}\\
{\itshape RE: CALIB\textsc{ratio}N OF CAMERAS London}\\
wassup fish?\\

@@

\rule{0.8\textwidth}{.4pt}
\clearpage
{\bfshape 2000-02-26 20:41:20 Trash}\\
{\itshape RE: teaching / TRIBES}	\\
Basically think about it like this... On the Earth there are many different languages spoken by people usually in a specific geographical location. Now myself and yeahbaby looked at this experiment (as we are very interested in AI and have done much work in this field) and decided to test the idea of ``Tribes" where all those in that tribe speak the same lingo. We also figured that tribes rarely move out of their ``Birth Location" so kept them at one site. As can be seen now ALL our Agents speak the same language and have even taught it to others outside the Tribe (such as the esteemed Yellow Mk 1 Escort  and possibly even Anne's little Agent). 
These agents we considert to be forreigners!! But stilll they can communicate with our Agents... If you have noticed our words in the lexicon as explained by yeahbaby have  very few deviations from what was intended unlike others such as ``Gumble" which appears to mean Everything!!!!!

@@

{\bfshape 2000-02-27 19:52:20	Trash}\\
{\itshape RE: Collusion Tribes}\\
At the minute me and Yeah are off the Paris server (we are regrouping our thoughts for further action!!!)
But yes I did mean view the pictures when i said definition.\\
By 2 tribes I mean obviously two who have been taught on the same server.
So say that i train my agents on there this week and you train your tribe next week on the same server with different words..... when both tribes are capable of good communication we then take half of each tribe and place them together in the 3rd week on the Paris server and see if they become multi lingual.\\
We then sned these linguists back to their own tribe and see what happens.\\ \clearpage
Unfortunately many people have tried to jump on the band wagon of Paris success and send their Agents there with no idea what they are doing ... this has meant that we can not get our tribe to dominate as we would like.....\\
The problems with organising tribe transfer is immense and it is a shame people don't exp[lain what they are doing on this fprum so that all could help out by say teaching their agents the words being used or even evacuating a site so that two tribes can collide ...\\
It is a real shame that an experiment into AI and Language is hampered because people are unwilling to communicate with each other .... \\
Well I guess that those of us who want a proper experiment should just ry their best in difficult circumstances...\\
About the wrong words being taught !!! If your Tribe is more than 60\% of the server then your tribe generallly succeeds in teaching the ``foreigners'' with little effect on themselves ... This has more to do with Probability than anything else.

@@

{\bfshape 2000-03-09 15:52:11 Trash}\\
{\itshape RE: Yeah how about it!!}\\
It matters not who screws who at thsio point the fact remains if they want to run this experiment their way and by their rules they should set it up purely in their own labs .... without us having access if they want us to contribute then they should let us do it our way ..\\
Hey Luc and all the rest of you get on this forum and have the balls to talk to us"\\

@@

``Look into the box" consisted of a box in which a camera was mounted that would take a picture of the 
eye of a person who looked into the box. The eye was then projected much bigger on a wall opposite to the box, 
which in itself gave a very strong visual effect (see \figref{fig:lookintobox}). 
At the same time two artificial agents looked at the 
eye colour and played a colour language game. Visitors to the exhibition could hear the dialog between the agents 
and follow on a nearby screen the progression in the emergence of a colour vocabulary. During the course of the 
exhibition an artificial colour language emerged, reflecting the eye and skin colour of visitors.

@@

The Look into the Box installation was also shown during the ``Intensive Science" exhibitions organised in October 
2006 on the occasion of the 10th anniversary of the Sony Computer Science Laboratory.\is{Intensive Science exhibition}
It was first shown in Paris at 
the exhibition space ``La Maison Rouge" near the Bastille in Paris, which featured science/art installations 
and collaborations involving members of the Sony Computer Science Laboratory, including work by Atau Tanaka, 
Francois Pachet in collaboration with Jazz pianist Albert van Veendaal, Peter Hanappe in collaboration with 
photographer Armin Linke, Frederic Kaplan in collaboration with the Design School of EPFL in Lausanne. 
The same exhibition traveled to Tokyo where it was shown in the Sony Explorascience museum from 22 December 2006 
to 12 February 2007 (see \figref{fig:intensive-science}). 
\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap9/figs/explorascience.pdf}}
\caption{\label{fig:intensive-science}Installation at the Tokyo Explorascience Museum. The box is in the left corner. The eye is projected on a 
big screen. A smaller screen shows the image captured by the camera, the colours recognised, and the words used by 
the agents.}
\end{figure}
\section{Conclusions}

@@

The installations were also one of the pioneering attempts to develop a strong interaction between art and science, 
which today are more common and more recognised than they were in the late 90's.
They were conceived to be part of exhibitions and well established artists, such as Olafur Eliasson, took great 
interest and collaborated to give them a twist to function better within an art context. 
The context of an exhibition is a very effective way to reach an audience
that has normally no access to ongoing scientific experiments. It also generated a stream of newspaper articles, 
and contributions to radio and television programs, which normally pay only little attention to what goes in science.

>>>>

\vspace{1\baselineskip}
\noindent{\bfshape Situated robots and teleports}

@@

\vspace{1\baselineskip}
\noindent{\bfshape Constructing perceptually grounded concepts}

@@

\vspace{1\baselineskip}
\noindent{\bfshape Sources of noise}

@@

\vspace{1\baselineskip}
\noindent{\bfshape Cultural evolution}

@@

\clearpage We pick up the dialog on the 25 February 2000. Until that date the forum was very active both 
with general discussions about the future of intelligence or the origins of language, and with specific questions, mostly 
why agents were so slow in playing games. Many users were too impatient and did not seem to realise that playing 
a language game could easily take a minute or two. However the major problem was the deviating alignment of the cameras, 
which had to guarantee that there was a common frame of reference between the agents, and hence the 
possibility of sharing attention. The cameras were mounted on tripods. When somebody accidentally 
moved the tripod, the frame of reference was no longer exact. As mentioned earlier, traffic (particularly subway 
traffic in London) caused vibrations of the cameras so that they kept shifting and getting out of balance.

@@

{\bfshape 2000-03-03 20:08:36 schedski}\\
{\itshape RE: CALIB\textsc{ratio}N OF CAMERAS London}\\
wassup fish?\enlargethispage{1\baselineskip}\\

@@

{\bfshape 2000-02-26 20:41:20 Trash}\\
{\itshape RE: teaching / TRIBES}	\\
Basically think about it like this... On the Earth there are many different languages spoken by people usually in a specific geographical location. Now myself and yeahbaby looked at this experiment (as we are very interested in AI and have done much work in this field) and decided to test the idea of ``Tribes" where all those in that tribe speak the same lingo. We also figured that tribes rarely move out of their ``Birth Location" so kept them at one site. As can be seen now ALL our Agents speak the same language and have even taught it to others outside the Tribe (such as the esteemed Yellow Mk 1 Escort  and possibly even Anne's little Agent). 
These agents we considert to be forreigners!! But stilll they can communicate with our Agents... If you have noticed our words in the lexicon as explained by yeahbaby have  very few deviations from what was intended unlike others such as ``Gumble" which appears to mean Everything!!!!!

@@

{\bfshape 2000-02-27 19:52:20	Trash}\\
{\itshape RE: Collusion Tribes}\\
At the minute me and Yeah are off the Paris server (we are regrouping our thoughts for further action!!!)
But yes I did mean view the pictures when i said definition.\\
By 2 tribes I mean obviously two who have been taught on the same server.
So say that i train my agents on there this week and you train your tribe next week on the same server with different words..... when both tribes are capable of good communication we then take half of each tribe and place them together in the 3rd week on the Paris server and see if they become multi lingual.\\
We then sned these linguists back to their own tribe and see what happens.\\
Unfortunately many people have tried to jump on the band wagon of Paris success and send their Agents there with no idea what they are doing ... this has meant that we can not get our tribe to dominate as we would like.....\\
The problems with organising tribe transfer is immense and it is a shame people don't exp[lain what they are doing on this fprum so that all could help out by say teaching their agents the words being used or even evacuating a site so that two tribes can collide ...\\
It is a real shame that an experiment into AI and Language is hampered because people are unwilling to communicate with each other .... \\
Well I guess that those of us who want a proper experiment should just ry their best in difficult circumstances...\\
About the wrong words being taught !!! If your Tribe is more than 60\% of the server then your tribe generallly succeeds in teaching the ``foreigners'' with little effect on themselves ... This has more to do with Probability than anything else.

@@

{\bfshape 2000-03-09 15:52:11 Trash}\\
{\itshape RE: Yeah how about it!!}\\
It matters not who screws who at thsio point the fact remains if they want to run this experiment their way and by their rules they should set it up purely in their own labs .... without us having access if they want us to contribute then they should let us do it our way ..\\
Hey Luc and all the rest of you get on this forum and have the balls to talk to us"\enlargethispage{1\baselineskip}\\

@@

``Look into the Box" consisted of a box in which a camera was mounted that would take a picture of the 
eye of a person who looked into the box. The eye was then projected much bigger on a wall opposite to the box, 
which in itself gave a very strong visual effect (see \figref{fig:lookintobox}). 
At the same time two artificial agents looked at the 
eye colour and played a colour language game. Visitors to the exhibition could hear the dialog between the agents 
and follow on a nearby screen the progression in the emergence of a colour vocabulary. During the course of the 
exhibition an artificial colour language emerged, reflecting the eye and skin colour of visitors.

@@

\clearpage The Look into the Box installation was also shown during the ``Intensive Science" exhibitions organised in October 
2006 on the occasion of the 10th anniversary of the Sony Computer Science Laboratory.\is{Intensive Science exhibition}
It was first shown in Paris at 
the exhibition space ``La Maison Rouge" near the Bastille in Paris, which featured science/art installations 
and collaborations involving members of the Sony Computer Science Laboratory, including work by Atau Tanaka, 
Francois Pachet in collaboration with Jazz pianist Albert van Veendaal, Peter Hanappe in collaboration with 
photographer Armin Linke, Frederic Kaplan in collaboration with the Design School of EPFL in Lausanne. 
The same exhibition traveled to Tokyo where it was shown in the Sony Explorascience museum from 22 December 2006 
to 12 February 2007 (see \figref{fig:intensive-science}). 
\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap9/figs/explorascience.pdf}}
\caption{\label{fig:intensive-science}Installation at the Tokyo Explorascience Museum. The box is in the left corner. The eye is projected on a 
big screen. A smaller screen shows the image captured by the camera, the colours recognised, and the words used by 
the agents.}
\end{figure}
\section{Conclusions}

@@

The installations were also one of the pioneering attempts to develop a strong interaction between art and science, 
which today are more common and more recognised than they were in the late 1990s.
They were conceived to be part of exhibitions and well established artists, such as Olafur Eliasson, took great 
interest and collaborated to give them a twist to function better within an art context. 
The context of an exhibition is a very effective way to reach an audience
that has normally no access to ongoing scientific experiments. It also generated a stream of newspaper articles, 
and contributions to radio and television programs, which normally pay only little attention to what goes in science.
